Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-13 main.tex
Cp experiments with definitions

@@

Intuitively, we use the word {\em policy} for any kind of dynamic
enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
capability hardware, etc.). We use {\em property} for a mathematical
predicate on programs (i.e., machine states) articulating some useful
security goal, e.g. noninterference.

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}\times
\{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
levels,'' where $\HIGH$ marks locations containing secret or low-integrity
data and $\LOW$ marks those with public or high-integrity data, and
``observabilities,'' where $\OBS$ marks locations whose contents are
observable from the outside (by the ``attacker'') and $\HIDDEN$ marks
internal ``hidden'' state that cannot be observed directly.

@@

\bcp{Sketchy / out of date from here on...}

  \begin{itemize}
  \item A {\em property} is a function that takes a contour to a
  prefix-closed set of (finite and/or infinite) sequences of machine states.
  That is, it takes a starting contour and produces a checker that tells
  whether a given trace is OK to observe.
  \begin{itemize}
  \item I am not sure whether we want to be talking in terms of {\em
    properties} or {\em hyperproperties} here (and all the way through).
  \end{itemize}
  \item Proposal: Forget about full-blown NI and focus the whole paper on
  variants of ``lockstep NI'' where the PC is the same in every pair of
  related runs.
  \begin{itemize}
  \item One issue that I am not sure about: Could there be a sneaky way of
  ``putting secret information into the control state'' where the PCs of
  related runs are always in lockstep but, because the instruction memory is
  writeable, different instructions can get dispatched.

  \item So maybe we should be asking that the instruction stream be
  identical, not (just?) the PCs
  \end{itemize}
  \item One nice observation is that, if we are only interested in pairs of
  runs with identical PCs, there are two different ways to proceed: (1) say
  that we {\em ignore} runs where the PCs (or instruction streams) diverge
  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
  instruction streams) diverge, i.e., demand that this never happens (e.g.,
  stack and heap policies).
  \item A remaining question is what to do about the possibility that the
  contour changes differently on two related runs.  We could disallow this,
  but not certain it's necessary---maybe everything works out OK.
  \end{itemize}

@@

Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
every pair $(\mach,\mach')$ of variant machine states with respect to
$\contour$ are indistinguishable under $\pol$.

@@

\item Relation to the standard definition of NI.

\item Fact: The standard RISC-V semantics does not guarantee NI.  (Give a
little example of a program that reads from a secret location and writes to
a public one, just to illustrate the definition of NI.)
\end{itemize}
\fi

>>>>

\subsection{Contours and Variants}

Intuitively, we use the word {\em policy} for any kind of dynamic
enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
capability hardware, etc.). We use {\em property} for a mathematical
predicate on programs (i.e., machine states) articulating some useful
security goal, e.g. noninterference. \iftext\bcp{Not quite true, if we stick
  with hyperproperties.} \fi

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}\times
\{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
levels,'' where $\HIGH$ marks locations that may contain secret or
low-integrity data and $\LOW$ marks those that must only ever contain
publicly readable or high-integrity data, and ``observabilities,'' where
$\OBS$ marks locations whose contents are observable from the outside (by
the ``attacker,'' e.g., by DMA) and $\HIDDEN$ marks internal ``hidden''
state that cannot be observed directly (e.g., processor registers and memory
addresses for temporary variables will usually be hidden).

@@

@@

\subsection{Properties}

\bcp{Here is where it starts getting iffy...}

A {\em contoured (hyper)property} is a function that takes a contour to a
hyperproperty.  That is, it takes a starting contour and produces a
\bcp{prefix-closed?} set of pairs of sequences of machine states.

@@

@@

\item Relation to the standard definition of NI.
\end{itemize}
\fi

\bcp{Recheck.}
Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
every pair $(\mach,\mach')$ of variant machine states with respect to
$\contour$ are indistinguishable under $\pol$.

\iftext
\bcp{
Fact: The standard RISC-V semantics (i.e., the null policy) does not
guarantee NI.  (Give a little example of a program that reads from a secret
location and writes to a public one, just to illustrate the definition of
NI.)}
\fi

\subsection{More...}
  \begin{itemize}
  \item Proposal: Maybe forget about full-blown NI and focus the whole paper
  on variants of ``lockstep NI'' where the PC is the same in every pair of
  related runs?
  \item One issue that I am not sure about: Could there be a sneaky way of
  ``putting secret information into the control state'' where the PCs of
  related runs are always in lockstep but, because the instruction memory is
  writeable, different instructions can get dispatched.

  So maybe we should be asking that the instruction stream be
  identical, or maybe instructions paired with PCs?

  \item One nice observation is that, if we are only interested in pairs of
  runs with identical PCs, there are two different ways to proceed: (1) say
  that we {\em ignore} runs where the PCs (or instruction streams) diverge
  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
  instruction streams) diverge, i.e., demand that this never happens (e.g.,
  stack and heap policies).
  \item A remaining question is what to do about the possibility that the
  contour changes differently on two related runs.  We could disallow this,
  but not certain it's necessary---maybe everything works out OK.
  \end{itemize}
